A Sound and Complete Deep Embedding of Arbitrary Normal Modal Logics in Coq

### 2025-08-20 paper/COLA/paper.tex
Future work: soundness for other systems
Clarify completeness result

The proof proceeds by induction on the derivation of $\Gamma \vdash\varphi$ and
it has a case for each of the axioms of the Hilbert system for propositional
logics, one for each axioms $K$ and possibility, and the rules for modus
ponens and necessitation. In order to modularize the proof, we defined a lemma for each case.
As example, the following is the statement for modus ponens case.
\begin{minted}{coq}
Lemma Modus_Ponens_soundness:
  forall M w p q, (M ' w ||- p) /\ (M ' w ||- [! p -> q !]) -> (M ' w ||- q).
\end{minted}
All these lemmas are direct consequences from the definitions of the Kripke
semantics and entailment. We note that system K is used as it doesn't impose
any restrictions on the frame:
i.e., the relation on the frame may be an arbitrary relation.

@@

In this section, we describe our formalization of the completeness result for the
Hilbert proof system with respect to the Kripke semantics.
Before we start,
it is worth mentioning that there are different notions of completeness, and
the literature is not always clear about the difference of those~\cite{Popkorn_1994,modal_lean}.
%-------------------------------------------------------------------------------
The semantic entailment, $\Gamma\models\varphi$, represents a global consequence since, by our definition,
we expect that the hypotheses are valid in every world in a model instead
of having $\varphi$ hold only on those worlds where the hypotheses are valid \cite{de1999global}.
Also, the completeness result may be
weak, which is only valid without hypotheses, or strong, allowing it to be stated with any set of
hypotheses \cite{blackburn}. We follow the definitions as stabilished by Popkorn \cite{Popkorn_1994}
and note that we prove a strong completeness result using a global consequence: as this subsumes
both weak completeness and a local one (as local consequence implies global consequence
\cite{de1999global}), our result allows us to derive other possible definitions of completeness,
while still being a weak result enough that it will work for any of the possible sets of
axioms \cite{Popkorn_1994}.

>>>>

The proof proceeds by induction on the derivation of $\Gamma \vdash\varphi$ and
it has a case for each of the axioms of the Hilbert system for propositional
logics, one for each axioms $K$ and possibility, and the rules for modus
ponens and necessitation. In order to modularize the proof, we defined a lemma for each case.
As example, the following is the statement for modus ponens case.
\begin{minted}{coq}
Lemma Modus_Ponens_soundness:
  forall M w p q, (M ' w ||- p) /\ (M ' w ||- [! p -> q !]) -> (M ' w ||- q).
\end{minted}
All these lemmas are direct consequences from the definitions of the Kripke
semantics and entailment. We note that system K is used as it doesn't impose
any restrictions on the frame:
i.e., the relation on the frame may be an arbitrary relation.
We intend to cover soundness for larger systems as future work, which requires
checking the additional axioms individually.

@@

In this section, we describe our formalization of the completeness result for the
Hilbert proof system with respect to the Kripke semantics.
Before we start,
it is worth mentioning that there are different notions of completeness, and
the literature is not always clear about the difference of those~\cite{Popkorn_1994,modal_lean}.
%-------------------------------------------------------------------------------
The semantic entailment, $\Gamma\models\varphi$, represents a global consequence since, by our definition,
we expect that the hypotheses are valid in every world in a model instead
of having $\varphi$ hold only on those worlds where the hypotheses are valid \cite{de1999global}.
Also, the completeness result may be
weak, which is only valid without hypotheses, or strong, allowing it to be stated with any set of
hypotheses \cite{blackburn}. We follow the definitions as established by Popkorn \cite{Popkorn_1994}
and note that we prove a weak completeness result using a global consequence, which subsumes a local one (as local consequence implies global consequence
\cite{de1999global}). A strong completeness proof, as of now, is left as future work.
It's worth noting that our completeness theorem is, in the words of
Popkorn, \emph{rather weak} \cite[p. 119-120]{Popkorn_1994}; the result is applicable to any
possible sets of axioms extending System K. Other
notions of completeness, argueable more expressive, will relate sets of axioms
to respective sets of model frames (cf. Section \ref{sec:correspondence}). Still, the present result provides a basis
for most such stronger completeness theorems.
